U.S. Labour Market Resilience

### 2024-03-18 paper/paper.qmd
Update paper

This paper uses U.S. General Social Survey (GSS) [@GSS] data to examine the "number of hours worked last week" as a key measure of labour engagement. We hypothesize that economic downturns lead to decreased labour engagement, seen in reduced work hours, impacting labour groups differently. By comparing periods of economic crises with times of stability, we aim to understand the labour market's responses, focusing on the most affected sectors and demographic groups.

@@

Initial findings show disparities: workers faced significant reductions in working hours during the crisis, with studies by [@Chen_Yeh_2021] and [@Singh_Shirazi_Turetken_2022] showing that women face a more severe job impact due to their unique social obligations. This highlights the need for targeted policy interventions to address these differences and enhance labour market resilience.

@@

The paper uses data collected from the US General Social Survey (GSS) from NORIC at the University of Chicago [@GSS]. From the dataset, this paper focuses on the "number of hours worked last week" variable, from the years 1972 to 2022. This longitudinal approach allows us to compare labour dynamics across different economic downturns, including the 1998 Asian Financial Crisis, the 2008 Global Financial Crisis, and the COVID-19 pandemic in 2020.

@@

From 1972 to 2018, the GSS was conducted through face-to-face interviews. In-person interviews facilitated a richer data collection environment, where interviewers could clarify ambiguities and probe deeper into responses. This mode of data collection was ideal and allowed interviewers to collect richer data. However, in response to the COVID-19 Pandemic, from 2020 to 2021, the GSS transitioned to web surveys as the primary method of collecting data.

@@

The survey's methodology has several strengths. From 1972 to 2018, data was gathered through face-to-face interviews, enabling interviewers to capture diverse and comprehensive information, providing a deeper understanding of respondents. Another notable strength is the survey's consistency. The General Social Survey (GSS) has maintained a uniform approach since 1972, with minimal alterations in questions and data collection methods each year. This consistency facilitates clear year-to-year analyses for researchers. The survey, traditionally administered through face-to-face interviews, transitioned to web surveys during the COVID-19 Pandemic. Relevance is another key strength of the GSS. It encompasses a broad spectrum of demographic, behavioural, and attitudinal questions, covering topics such as civil liberties, crime and violence, and psychological well-being among individuals in the United States [@gssCodebook2022]. With its almost annual frequency, the survey delivers up-to-date and pertinent information across various fields. Furthermore, the survey promotes accessibility by being available to the public. This means that anyone interested in the survey's results can freely access the data, enhancing transparency and openness in disseminating valuable information.

@@

Since 1972, the General Social Survey (GSS) has been gathering data primarily through face-to-face interviews, during which participants were interviewed using a structured questionnaire. This approach persisted until the onset of the COVID-19 Global Pandemic, which resulted in the suspension of data collection for the year 2020. Subsequently, the data collection methodology underwent a transition from in-person interviews to web-based questionnaires. This pandemic-induced shift introduced several variables that could potentially impact data quality and comparability. This alteration raises concerns and stands as a notable limitation on the accuracy of reported work hours and the perceived effects of economic crises. The move from traditional in-person interviews to web-based surveys brings uncertainties, as the absence of interactive guidance may lead to misinterpretations or superficial engagement with survey questions. An essential drawback in the dataset is the limited number of questions asked. Only a single survey question was asked, focusing only on obtaining numerical data regarding the respondent's worked hours. This weakness limits the understanding of the respondent's demographic details, posing a challenge in the analysis process.

@@

However, with the introduction of web-based surveys, new categories like "Skipped on Web" were introduced to categorize non-responses. This change raises questions about the comparability of data across years and the potential for increased non-response rates. Non-responses can significantly affect the analysis of labour market trends, particularly in capturing the full extent of unemployment or underemployment during economic downturns. The handling of these non-responses requires scrutiny to assess how it might influence findings on work hours and job stability across different demographic groups and economic sectors.

@@

```{r}
#| label: fig-demographic-comparison
#| fig-cap: Working Hours Comparisons with Demographic factors
#| layout-ncol: 2
#| echo: false
#| warning: false
#| message: false

@@

@@

We agree that strong measures should be in place to ensure participant confidentiality and data protection, as mentioned in @gssCodebook. This involves using a secure database to store sensitive information and encrypting participant details to prevent unauthorized access.

In addition, we adopted several strategies to address potential bias in our research. Recognizing the risk of sampling bias, whereby certain groups may be over-represented, we have diversified our outreach methods. Inspired by @gssCodebook, we recommend that new survey approaches post-2020 could reach participants through a variety of channels, including social media, community communications, and local organizations to ensure a broad range of demographics.

@@

Our analysis grounded in the rich dataset provided by the General Social Survey (GSS) [^1], navigates through a complex terrain of labour market dynamics in the wake of economic crises. However, as mentioned in the data @sec-data, several limitations inherent to the dataset and the broader research approach necessitate a cautious interpretation of our findings.

The original data only provides simple data on working hours, but in @gssCodebook, GSS has a detailed explanation of the data survey method, such as interviewing random people through questionnaires. The sensitive nature of questions related to economic hardships and employment challenges may lead to a degree of response bias. Despite advancements in survey techniques since the 1970s, the discomfort associated with disclosing personal economic impacts or job losses can limit the diversity and authenticity of responses collected. Framing questions within the GSS regarding employment and economic resilience may also inadvertently reflect biases or presuppositions. This phenomenon could mirror previous findings where respondents mask their true beliefs on topics, potentially skewing the data towards more socially acceptable responses.

@@

Finally, our paper also has certain limitations in processing data as shown in data @sec-data. One limitation stems from our decision to analyze the years immediately following the economic crisis (2010 and 2021) rather than the year immediately following the recession itself. This choice was made partly because of the lack of data for 2009 in the General Social Survey (GSS), necessitating the need to use 2010 (the second year of the economic crisis) as a proxy to measure the post-crisis impact of the 2008 financial recession. Likewise, given the unprecedented impact of the COVID-19 pandemic in 2020, the selection of 2021 was intended to capture its consequences rather than 2022, which could be compared to the 2010 selection.

@@

In addressing non-response rates within labour market surveys, @Groves_Peytcheva_2008 provides some suggestions for directions for our consideration. They advocate for tailored communication and mixed-mode survey approaches as effective measures against non-response bias (Journal of Official Statistics, 24(2), 167-189). For instance, future research could look into the success of the Health and Retirement Study, which employs a mixed-mode design, combining in-person interviews with telephone and online follow-ups to accommodate participant preferences and improve response rates. These examples underscore the potential of adopting innovative methodologies and technologies in labour market surveys to ensure more representative data.

@@

To tackle response bias, a common hurdle in surveys on sensitive topics like economic impact, future efforts could lean more heavily into techniques that safeguard participant anonymity and confidentiality. For example, the Pew Research Center's [@Pew_Research_Center_2021] method of ensuring anonymity in their political opinion polls could serve as a model, enhancing respondent confidence and honesty. Briefly, their approach is characterized by random sampling to select representativeness. The main thing is that they will demonstrate to the public how the respondent's data is protected through encryption and strict privacy protocols, ensuring that no personally identifiable information is recorded or stored.

@@

For future labour market survey research, we refer to the method proposed by @Groves_Peytcheva_2008 above to eliminate non-response rates and bias as much as possible. We recommend a hybrid offline and online survey design; this approach can improve participation rates and data representativeness by accommodating a wider range of respondent preferences and accessibility. For example, incorporating new technologies such as mobile survey apps or secure online platforms could reduce participation barriers.

>>>>

This paper uses U.S. General Social Survey (GSS) data to examine the "number of hours worked last week" as a key measure of labour engagement. We hypothesize that economic downturns lead to decreased labour engagement, seen in reduced work hours, impacting labour groups differently. By comparing periods of economic crises with times of stability, we aim to understand the labour market's responses, focusing on the most affected sectors and demographic groups.

@@

Initial findings show disparities: workers faced significant reductions in working hours during the crisis. In 2021, more than 50% of workers worked 21-40 hours, and only more than 10% will work 41-60 hours; while in 2018, an average economic year, more than 40% of workers will work 41-60 hours. Furthermore, studies by @Chen_Yeh_2021 and @Singh_Shirazi_Turetken_2022 show that women face a more severe job impact due to their unique social obligations. This highlights the need for targeted policy interventions to address these differences and enhance labour market resilience.

@@

The paper uses data collected from the US General Social Survey (GSS) from NORIC at the University of Chicago [@GSS]. From the dataset, this paper focuses on the "number of hours worked last week" variable, from the years 1992 to 2022. This longitudinal approach allows us to compare labour dynamics across different economic downturns, including the 1998 Asian Financial Crisis, the 2008 Global Financial Crisis, and the COVID-19 pandemic in 2020.

@@

From 1992 to 2018, the GSS was conducted through face-to-face interviews. In-person interviews facilitated a richer data collection environment, where interviewers could clarify ambiguities and probe deeper into responses. This mode of data collection was ideal and allowed interviewers to collect richer data. However, in response to the COVID-19 Pandemic, from 2020 to 2021, the GSS transitioned to web surveys as the primary method of collecting data.

@@

The survey's methodology has several strengths. From 1992 to 2018, data was gathered through face-to-face interviews, enabling interviewers to capture diverse and comprehensive information, providing a deeper understanding of respondents. Another notable strength is the survey's consistency. The General Social Survey (GSS) has maintained a uniform approach since 1992, with minimal alterations in questions and data collection methods each year. This consistency facilitates clear year-to-year analyses for researchers. The survey, traditionally administered through face-to-face interviews, transitioned to web surveys during the COVID-19 Pandemic. Relevance is another key strength of the GSS. It encompasses a broad spectrum of demographic, behavioural, and attitudinal questions, covering topics such as civil liberties, crime and violence, and psychological well-being among individuals in the United States [@gssCodebook2022]. With its almost annual frequency, the survey delivers up-to-date and pertinent information across various fields. Furthermore, the survey promotes accessibility by being available to the public. This means that anyone interested in the survey's results can freely access the data, enhancing transparency and openness in disseminating valuable information.

@@

Since 1992, the General Social Survey (GSS) has been gathering data primarily through face-to-face interviews, during which participants were interviewed using a structured questionnaire. This approach persisted until the onset of the COVID-19 Global Pandemic, which resulted in the suspension of data collection for the year 2020. Subsequently, the data collection methodology underwent a transition from in-person interviews to web-based questionnaires. This pandemic-induced shift introduced several variables that could potentially impact data quality and comparability. This alteration raises concerns and stands as a notable limitation on the accuracy of reported work hours and the perceived effects of economic crises. The move from traditional in-person interviews to web-based surveys brings uncertainties, as the absence of interactive guidance may lead to misinterpretations or superficial engagement with survey questions. An essential drawback in the dataset is the limited number of questions asked. Only a single survey question was asked, focusing only on obtaining numerical data regarding the respondent's worked hours. This weakness limits the understanding of the respondent's demographic details, posing a challenge in the analysis process.

@@

However, with the introduction of web-based surveys, new categories like "Skipped on Web" were introduced to categorize non-responses. This change raises questions about the comparability of data across years and the potential for increased non-response rates. Non-responses (@fig-nonresponse-rate) can significantly affect the analysis of labour market trends, particularly in capturing the full extent of unemployment or underemployment during economic downturns. The handling of these non-responses requires scrutiny to assess how it might influence findings on work hours and job stability across different demographic groups and economic sectors.

@@

```
\break

The data depicted in @fig-hour-working-percentage illustrates the average yearly working hours. Initially, it is evident that the highest average yearly working hours fall within the range of 21-40 hours. Particularly notable are the years 2021 and 2022, which exhibit the highest percentage of average working hours. One potential explanation for this pattern could be linked to the post-pandemic economy, marked by efforts towards economic recovery and escalating costs across various industries. As more individuals reentered the workforce, it is plausible that this contributed to the elevated average working hours observed in 2021 and 2022.

@fig-hour-working-percentage indicates that the second most common range of average working hours is between 41-60 hours, suggesting that a significant portion of the population works within this time frame. Following this, the analysis highlights that the third most common range is 0-20 hours, indicating a substantial number of individuals who work part-time or have limited working hours. As the analysis progresses, it notes that the averages of 61-80 hours and 80+ hours represent progressively smaller percentages of average working hours. This implies that fewer individuals work such long hours compared to those in the previous ranges. Finally, it specifies that the category of 80+ hours represents the smallest percentage, indicating that only a minority of individuals work exceptionally long hours. Overall, this analysis provides insight into the distribution of working hours, highlighting common patterns and indicating trends towards fewer individuals working extremely long hours.

```{r}
#| label: fig-demographic-comparison
#| fig-cap: Working Hours Comparisons with Demographic factors
#| fig-subcap: ["Average Age by Each Year","Percentage of Female Respondent by Each Year"]
#| layout-ncol: 2
#| echo: false
#| warning: false
#| message: false

@@

average_age_by_year <- aggregate(AGE ~ YEAR, data = sup_data, FUN = mean)
sex_counts_by_year <- aggregate(SEX ~ YEAR, data = sup_data, function(x) mean(x == 2) * 100) #calculate percentage of females each year

average_age_by_year <- average_age_by_year |> 
  filter(YEAR%in%c(1998,2008, 2010,2018,2021))

sex_counts_by_year <- sex_counts_by_year |> 
  filter(YEAR%in%c(1998,2008,2010,2018,2021))

average_ages <- average_age_by_year$AGE
average_gender <- sex_counts_by_year$SEX

#Histogram of average working hours for 2008, 2021, 2022, and Total
hist_data <- select(data, work_hours,"1998","2008","2010","2018","2021")
hist_data <- hist_data[5:94,]

num <- 5 #number of years
sum <- rep(0,num)
total_people <- rep(0,num)
for(i in 1:num){
  total_people[i] <- sum(hist_data[, i+1])
}
hist_data[, 1] <- sapply(hist_data[, 1], as.numeric)
hist_data[90,1] <- 90

sum[1] <- sum(hist_data$"1998" * hist_data$work_hours)
sum[2] <- sum(hist_data$"2008" * hist_data$work_hours)
sum[3] <- sum(hist_data$"2010" * hist_data$work_hours)
sum[4] <- sum(hist_data$"2018" * hist_data$work_hours)
sum[5] <- sum(hist_data$"2021" * hist_data$work_hours)

averages <- round(sum/total_people) 
years <- c("1998","2008","2010","2018","2021")

gg_data <- cbind(average_hours,average_ages,average_gender)

ggplot(gg_data,aes(x=years,y=average_ages)) +
  geom_bar(stat="identity") + 
  theme_minimal() + # Make the theme neater
  labs(x = "Years", y = "Average Age") +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom")

ggplot(gg_data,aes(x=years,y=average_gender)) +
  geom_bar(stat="identity") + 
  theme_minimal() + # Make the theme neater
  labs(x = "Years", y = "Percentage of Females (%)") +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom")
```

@fig-demographic-comparison presents a comparison of average ages across different years, specifically for 1998, 2008, 2010, 2018, and 2021. The y-axis represents the average age, and the x-axis denotes the years. The bars show a slight decrease in the average age in 2021 compared to the previous years, suggesting a younger demographic in the most recent year displayed. It also illustrates the percentage of females within the data for the same years as shown in @fig-demographic-comparison. The y-axis indicates the percentage of females, while the x-axis represents the years. The bars are relatively even, demonstrating a consistent representation of females across the selected years, with a very slight variation.

Both figures are titled "Working Hours Comparisons with Demographic factors," implying that the data presented may be used to analyze the relationship between demographic characteristics (age and gender) and working hours, although the specific hours worked are not shown in these charts. The consistent layout and design of the charts facilitate a clear and direct comparison of demographic trends over the years.

@@

We agree that strong measures should be in place to ensure participant confidentiality and data protection, as mentioned in the @gssCodebook. This involves using a secure database to store sensitive information and encrypting participant details to prevent unauthorized access.

In addition, we adopted several strategies to address potential bias in our research. Recognizing the risk of sampling bias, whereby certain groups may be over-represented, we have diversified our outreach methods. Inspired by the @gssCodebook, we recommend that new survey approaches post-2020 could reach participants through a variety of channels, including social media, community communications, and local organizations to ensure a broad range of demographics.

@@

Our analysis grounded in the rich dataset provided by the General Social Survey (GSS) [^1], navigates through a complex terrain of labour market dynamics in the wake of economic crises. However, as mentioned in the @sec-data section, several limitations inherent to the dataset and the broader research approach necessitate a cautious interpretation of our findings.

The original data only provides simple data on working hours, but in [@gssCodebook], GSS has a detailed explanation of the data survey method, such as interviewing random people through questionnaires. The sensitive nature of questions related to economic hardships and employment challenges may lead to a degree of response bias. Despite advancements in survey techniques since the 1970s, the discomfort associated with disclosing personal economic impacts or job losses can limit the diversity and authenticity of responses collected. Framing questions within the GSS regarding employment and economic resilience may also inadvertently reflect biases or presuppositions. This phenomenon could mirror previous findings where respondents mask their true beliefs on topics, potentially skewing the data towards more socially acceptable responses.

@@

Finally, our paper also has certain limitations in processing data as shown in @sec-data section. One limitation stems from our decision to analyze the years immediately following the economic crisis (2010 and 2021) rather than the year immediately following the recession itself. This choice was made partly because of the lack of data for 2009 in the General Social Survey (GSS), necessitating the need to use 2010 (the second year of the economic crisis) as a proxy to measure the post-crisis impact of the 2008 financial recession. Likewise, given the unprecedented impact of the COVID-19 pandemic in 2020, the selection of 2021 was intended to capture its consequences rather than 2022, which could be compared to the 2010 selection.

@@

In addressing non-response rates within labour market surveys, @Groves_Peytcheva_2008 provides some suggestions for directions for our consideration. They advocate for tailored communication and mixed-mode survey approaches as effective measures against non-response bias [@Groves_Peytcheva_2008]. For instance, future research could look into the success of the Health and Retirement Study, which employs a mixed-mode design, combining in-person interviews with telephone and online follow-ups to accommodate participant preferences and improve response rates. These examples underscore the potential of adopting innovative methodologies and technologies in labour market surveys to ensure more representative data.

@@

To tackle response bias, a common hurdle in surveys on sensitive topics like economic impact, future efforts could lean more heavily into techniques that safeguard participant anonymity and confidentiality. For example, the @Pew_Research_Center_2021's research method of ensuring anonymity in their political opinion polls could serve as a model, enhancing respondent confidence and honesty. Briefly, their approach is characterized by random sampling to select representativeness. The main thing is that they will demonstrate to the public how the respondent's data is protected through encryption and strict privacy protocols, ensuring that no personally identifiable information is recorded or stored.

@@

For future labour market survey research, we refer to the method proposed by Groves and Peytcheva (2008) above to eliminate non-response rates and bias as much as possible. We recommend a hybrid offline and online survey design; this approach can improve participation rates and data representativeness by accommodating a wider range of respondent preferences and accessibility. For example, incorporating new technologies such as mobile survey apps or secure online platforms could reduce participation barriers.
